Guaranteeing Safety and Liveness of Unmanned Aerial Vehicle Platoons on Air Highways

### 2016-05-09 docs/AIAA 2016 highway platooning/high_form.tex
Made the "easy" changes to AIAA paper according to Claire's comments
Made some modifications according to Jaime's comments; still need to incorporate the rest of Jaime's comments

% !TEX root = main.tex
\section{Air Highways}
We consider air highways to be virtual highways in the airspace on which a number of UAV platoons may be present. UAVs seek to arrive at some desired destination starting from their origin by traveling along a sequence of air highways. Air highways are intended to be the common pathways for many UAV platoons, whose members may have different origins and destinations. By routing platoons of UAVs onto a few common pathways, the air space becomes more tractable to analyze and intuitive to monitor. The concept of platoons will be proposed in Section \ref{sec:platooning}; for now, we focus on air highways in this section.

Let an air highway be denoted by the function $\hw: [0, 1] \rightarrow \mathbb{R}^2$. Such a highway lies in a horizontal plane of fixed altitude, with start and end points given by $\hw(0)\in\mathbb{R}^2$ and $\hw(1)\in\mathbb{R}^2$ respectively. For simplicity, we assume that the highway segment is a straight line segment, and the parameter $s$ indicates the position in some fixed altitude as follows: $\hw(s) = \hw(0) + s(\hw(1) - \hw(0))$. To each highway, we assign a speed of travel $v_\hw$ and specify the direction of travel to be the direction from $\hw(0)$ to $\hw(1)$, denoted using a unit vector $\hwd = \frac{\hw(1) - \hw(0)}{\lVert\hw(1) - \hw(0)\rVert_2}$.

@@

\subsection{The Air Highway Placement Problem}
Consider a map $\cmap:\mathbb{R}^2 \rightarrow \mathbb{R}$ which defines the cost $\cmap(\pos)$ incurred when a UAV flies over the position $\pos=(\pos_x,\pos_y)\in\mathbb{R}^2$. Given any position $\pos$, a large value of $\cmap(\pos)$ indicates that the position $\pos$ is costly or undesirable for a UAV to fly over. Locations with high cost could, for example, include densely populated areas and areas around airports. In general, the cost map $\cmap(\cdot)$ may be used to model cost of interference with commercial air spaces, cost of accidents, cost of noise pollution, etc., and can be designed by government regulation bodies.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{"fig/highway_illustration"}
	\caption{Highway illustration}
	\label{fig:hw_ill}
\end{figure}

>>>>

% !TEX root = main.tex
\section{Air Highways}
We consider air highways to be virtual highways in the airspace on which a number of UAV platoons may be present. UAVs seek to arrive at some desired destination starting from their origin by traveling along a sequence of air highways. Air highways are intended to be the common pathways for many UAV platoons, whose members may have different origins and destinations. By routing platoons of UAVs onto a few common pathways, the airspace becomes more tractable to analyze and intuitive to monitor. The concept of platoons will be proposed in Section \ref{sec:platooning}; in this section, we focus on air highways.

Let an air highway be denoted by the continuous function $\hw: [0, 1] \rightarrow \mathbb{R}^2$. Such a highway lies in a horizontal plane of fixed altitude, with start and end points given by $\hw(0)\in\mathbb{R}^2$ and $\hw(1)\in\mathbb{R}^2$ respectively. For simplicity, we assume that the highway segment is a straight line segment, and the parameter $s$ indicates the position in some fixed altitude as follows: $\hw(s) = \hw(0) + s(\hw(1) - \hw(0))$. To each highway, we assign a speed of travel $v_\hw$ and specify the direction of travel to be the direction from $\hw(0)$ to $\hw(1)$, denoted using a unit vector $\hwd = \frac{\hw(1) - \hw(0)}{\lVert\hw(1) - \hw(0)\rVert_2}$.

@@

\subsection{The Air Highway Placement Problem}
Consider a map $\cmap:\mathbb{R}^2 \rightarrow \mathbb{R}$ which defines the cost $\cmap(\pos)$ incurred when a UAV flies over the position $\pos=(\pos_x,\pos_y)\in\mathbb{R}^2$. Given any position $\pos$, a large value of $\cmap(\pos)$ indicates that the position $\pos$ is costly or undesirable for a UAV to fly over. Locations with high cost could, for example, include densely populated areas and areas around airports. In general, the cost map $\cmap(\cdot)$ may be used to model cost of interference with commercial airspaces, cost of accidents, cost of noise pollution, etc., and can be designed by government regulation bodies.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.95\textwidth]{"fig/highway_illustration"}
	\caption{Highway illustration}
	\label{fig:hw_ill}
\end{figure}

### 2016-05-09 docs/AIAA 2016 highway platooning/introduction.tex
Made some modifications according to Jaime's comments; still need to incorporate the rest of Jaime's comments

Optimal control and game theory are powerful tools for providing liveness and safety guarantees to controlled dynamical systems under bounded disturbances, and various formulations \cite{Bokanowski10,Mitchell05,Barron89} have been successfully used to analyze problems involving small numbers of vehicles \cite{Fisac15,Chen14,Ding08}. These formulations are based on Hamilton-Jacobi (HJ) reachability, which computes the reachable set, defined as the set of states from which a system is guaranteed to have a control strategy to reach a target set of states. Reachability is a powerful tool because reachable sets can be used for synthesizing both controllers that steer the system towards a set of goal states (liveness controllers), and controllers that steer the system away from a set of unsafe states (safety controllers). Furthermore, the HJ formulations are flexible in terms of system dynamics, enabling the analysis of non-linear systems. The power and success of HJ reachability analysis in previous applications cannot be denied, especially since numerical tools are readily available to solve the associated HJ Partial Differential Equation (PDE) \cite{LSToolbox,Osher02,Sethian96}. However, the computation is done on a grid, making the problem complexity scale exponentially with the number of states, and therefore with the number of vehicles. This makes the computation intractable for large numbers of vehicles.

In order to accommodate potentially thousands of vehicles simultaneously flying in the air, additional structure is needed to allow for tractable analysis and intuitive monitoring. An air highway system on which platoons of vehicles travel accomplishes both goals. However, many details of such a concept need to be addressed. Due to the flexibility of placing air highways compared to building ground highways in terms of highway location, even the problem of air highway placement can be a daunting task. To address this, in the first part of this paper, we propose a flexible and computationally efficient way based on \cite{Sethian96} to perform optimal air highway placement given an arbitrary cost map that captures the desirability of having UAVs fly over any geographical location. We demonstrate our method using the San Francisco Bay Area as an example. Once air highways are in place, platoons of UAVs can then fly in fixed formations along the highway to get from origin to destination.

>>>>

Optimal control and game theory are powerful tools for providing liveness and safety guarantees to controlled dynamical systems under bounded disturbances, and various formulations \cite{Bokanowski10,Mitchell05,Barron89} have been successfully used to analyze problems involving small numbers of vehicles \cite{Fisac15,Chen14,Ding08}. These formulations are based on Hamilton-Jacobi (HJ) reachability, which computes the reachable set, defined as the set of states from which a system is guaranteed to have a control strategy to reach a target set of states. Reachability is a powerful tool because reachable sets can be used for synthesizing both controllers that steer the system towards a set of goal states (liveness controllers), and controllers that steer the system away from a set of unsafe states (safety controllers). Furthermore, the HJ formulations are flexible in terms of system dynamics, enabling the analysis of non-linear systems. The power and success of HJ reachability analysis in previous applications cannot be denied, especially since numerical tools are readily available to solve the associated HJ partial differential equation (PDE) \cite{LSToolbox,Osher02,Sethian96}. However, the computation is done on a grid, making the problem complexity scale exponentially with the number of states, and therefore with the number of vehicles. This makes the computation intractable for large numbers of vehicles.

In order to accommodate potentially thousands of vehicles simultaneously flying in the air, additional structure is needed to allow for tractable analysis and intuitive monitoring by human beings. An air highway system on which platoons of vehicles travel accomplishes both goals. However, many details of such a concept need to be addressed. Due to the flexibility of placing air highways compared to building ground highways in terms of highway location, even the problem of air highway placement can be a daunting task. To address this, in the first part of this paper, we propose a flexible and computationally efficient method based on \cite{Sethian96} to perform optimal air highway placement given an arbitrary cost map that captures the desirability of having UAVs fly over any geographical location. We demonstrate our method using the San Francisco Bay Area as an example. Once air highways are in place, platoons of UAVs can then fly in fixed formations along the highway to get from origin to destination.

### 2016-05-09 docs/AIAA 2016 highway platooning/main.tex
Made the "easy" changes to AIAA paper according to Claire's comments
Made some modifications according to Jaime's comments; still need to incorporate the rest of Jaime's comments

\abstract{Recently, there has been immense interest in using unmanned aerial vehicles (UAVs) for civilian operations such as package delivery, aerial surveillance, and disaster response. As a result, UAV traffic management systems are needed to support potentially thousands of UAVs flying simultaneously in the airspace, in order to ensure their liveness and safety requirements are met. Currently, the analysis of large multi-agent systems cannot tractably provide these guarantees if the agents' set of maneuvers are unrestricted. In this paper, we propose to have platoons of UAVs flying on air highways in order to impose the air space structure that allows for tractable analysis and intuitive monitoring. For the air highway placement problem, we use the flexible and efficient fast marching method to solve the Eikonal equation, which produces a sequence of air highways that minimizes the cost of flying from an origin to any destination. Within the platoons that travel on the air highways, we model each vehicle as a hybrid system with modes corresponding to its role in the platoon. Using Hamilton-Jacobi reachability, we propose several liveness controllers and a safety controller that guarantee the success and safety of all mode transitions. For a single altitude range, our approach guarantees safety for one safety breach per vehicle; in the unlikely event of multiple safety breaches, safety can be guaranteed over multiple altitude ranges. We demonstrate the satisfaction of liveness and safety requirements through simulations of three common scenarios.}

@@

\section*{Nomenclature}
(Nomenclature entries should have the units identified)\\
\noindent\begin{tabular}{@{}lcl@{}}
$\cmap$ &=& Cost map \\
$\ccost$ &=& Cumulative cost of a path \\
$\ocost$ &=& Optimal cumulative cost of a path \\
$\hw$ &=& Air highway \\
$\hwd$ &=& Direction of travel of air highway \\
$\hws$ &=& A sequence of air highways \\
$\wpt$ &=& Waypoint \\
$x$ &=& System state (of a vehicle) \\
$\pos=(\pos_x, \pos_y)$ &=& Horizontal position \\
$\vel=(\vel_x, \vel_y)$ &=& Horizontal velocity \\
$\td$ &=& Time limit for descent during potential conflict \\
$\veh{i}$ &=& $i$th vehicle \\
$\vehSCS{i}$ &=& Set of vehicles for vehicle $i$ to consider for safety 
\end{tabular} \\

>>>>

\abstract{Recently, there has been immense interest in using unmanned aerial vehicles (UAVs) for civilian operations such as package delivery, aerial surveillance, and disaster response. As a result, UAV traffic management systems are needed to support potentially thousands of UAVs flying simultaneously in the airspace, in order to ensure their liveness and safety requirements are met. Currently, the analysis of large multi-agent systems cannot tractably provide these guarantees if the agents' set of maneuvers is unrestricted. In this paper, we propose to have platoons of UAVs flying on air highways in order to impose the airspace structure that allows for tractable analysis and intuitive monitoring. For the air highway placement problem, we use the flexible and efficient fast marching method to solve the Eikonal equation, which produces a sequence of air highways that minimizes the cost of flying from an origin to any destination. Within the platoons that travel on the air highways, we model each vehicle as a hybrid system with modes corresponding to its role in the platoon. Using Hamilton-Jacobi reachability, we propose several liveness controllers and a safety controller that guarantee the success and safety of all mode transitions. For a single altitude range, our approach guarantees safety for one safety breach per vehicle; in the unlikely event of multiple safety breaches, safety can be guaranteed over multiple altitude ranges. We demonstrate the satisfaction of liveness and safety requirements through simulations of three common scenarios.}

@@

\section*{Nomenclature}
\noindent\begin{tabular}{@{}lcl@{}}
$\cmap$ &=& Cost map \\
$\ccost$ &=& Cumulative cost of a path \\
$\ocost$ &=& Optimal cumulative cost of a path \\
$\hw$ &=& Air highway \\
$\hwd$ &=& Direction of travel of air highway \\
$\hws$ &=& A sequence of air highways \\
$\wpt$ &=& Waypoint \\
$x$ &=& System state (of a vehicle) \\
$\pos=(\pos_x, \pos_y)$ &=& Horizontal position \\
$\vel=(\vel_x, \vel_y)$ &=& Horizontal velocity \\
$\td$ &=& Time limit for descent during potential conflict \\
$\veh{i}$ &=& $i$th vehicle \\
$\vehSCS{i}$ &=& Set of vehicles for vehicle $i$ to consider for safety 
\end{tabular} \\
